---
#Reproducible Research: Peer Assessment 1#

=============================================================================================================
Load the libraries
```{r libraries, echo=TRUE}
library("ggplot2")
library("dplyr")
```

Load, read and transform the data
```{r readingdata, echo=TRUE}
setwd("//cib.net/shareparis/Home13/silvaol/PRIVE/COURSERA/Data Science R/Reproducible Research")
activity=read.csv("activity.csv")
```

**What is mean total number of steps taken per day?**

1. Calculate the total number of steps taken per day

```{r steps, echo=TRUE}
stepsbyday <- group_by(activity,date)
msteps<-summarize(stepsbyday,sum(steps,na.rm = TRUE))
names(msteps)<-(c("date","sum_steps"))
```

2. Make a histogram of the total number of steps taken each day
```{r hist, echo=TRUE}
ggplot(data=msteps, aes(msteps$sum_steps)) + geom_histogram(binwidth = 1500, col = "pink")+
xlab("Number of Steps") +
ylab("Count") +
ggtitle("Histogram of the Total Number of Steps Taken Each Day")
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r meanmedian, echo=TRUE}
mean(msteps$sum_steps)
median(msteps$sum_steps)
```

**What is the average daily activity pattern?**

1. Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r ts_plot, echo=TRUE}
five_min <- aggregate(steps ~ interval, data = activity, FUN =mean)
plot(five_min, type = "l")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r max_interval, echo=TRUE}
five_min[which.max(five_min$steps),]
```

**Imputing missing values**

1. Calculate and report the total number of missing values in the dataset
```{r total_NA, echo=TRUE}
apply(activity, 2, function(x) sum(is.na(x)))
```

2. Devise a strategy for filling in all of the missing values in the dataset.
```{r fill_NA, echo=TRUE}

replace_mean <- function(num) replace(num, is.na(num), mean(num, na.rm = TRUE))
activity %>% group_by(interval) %>% mutate(steps = replace_mean(steps))

```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r new_data, echo=TRUE}
new_data<-as.data.frame(activity %>% group_by(interval) %>% mutate(steps = replace_mean(steps)))
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

```{r hist2, echo=TRUE}
stepsbyday2 <- group_by(new_data,date)
msteps2<-summarize(stepsbyday2,sum(steps))
names(msteps2)<-(c("date","sum_steps2"))
ggplot(data=msteps2, aes(msteps2$sum_steps2)) + geom_histogram(binwidth = 1500, col = "green")+
xlab("Number of Steps") +
ylab("Count") +
ggtitle("Histogram of the Total Number of Steps Taken Each Day without NA")
```

```{r mean_median, echo=TRUE}
new_mean<-mean(msteps2$sum_steps2)
new_median<-median(msteps2$sum_steps2)
old_mean<-mean(msteps$sum_steps)
old_median<-median(msteps$sum_steps)
```

The first mean is `r old_mean`, the mean without NA is `r new_mean`
The first median is `r old_median`, the median without NA is `r new_median`

**Are there differences in activity patterns between weekdays and weekends?**

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r week_day, echo=TRUE}
new_data$day<-as.POSIXlt(new_data$date)$wday
new_data$Weekday <-ifelse(new_data$day > 5, "weekend", "weekday")
```

2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r final_plot, echo=TRUE}
five_min2 <- (new_data %>% group_by(interval, Weekday) %>% summarize(Mean = mean(steps)))
ggplot(five_min2, aes(interval, Mean)) + geom_line() + facet_grid(Weekday ~ .)+ylab("Mean of Steps") +
    ggtitle("Comparison of Average Number of Steps in Each Interval")
```
